Regression Cube Analysis of Cohort Study Data

### 2015-03-29 draft.tex
revised Application section

%\com{New Solution}.
%%
We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
%%
This approach combines all \emph{independent features} with respect to a \emph{target feature} and provides a visualization, which reveals insights into the data by highlighting relationships.
%%
A 3D-visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
%%
Slicing through the \emph{Regression Cube} allows for the detailed analysis of features towards the target disease.
%%
Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
%%
Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
%%
%\com{Validation \& Results}.
%%
We applied our \emph{Regression Cube} method to a hepatic steatosis data set to reproduce results from a data-mining driven analysis.
%%
A qualitative analysis with three domain experts was conducted on a breast fat data set.
%%
We were able to derive new hypotheses about relations between breast fat density and breast lesions towards breast cancer.
%%
%\com{Implications}.
%%
With the \emph{Regression Cube}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
} % end of abstract

@@

Testing features for association with diseases using regression analysis is one of the most important epidemiological tools.
%%
%Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the required subject count.
Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and required subject count.
%%
% missing overview techniques ist zu hart; es gibt ja techniken (eine scatterplot matrix zb.)
Due to the amount of data and only limited overview techniques, possible correlations may be missed.
%%
Explorative analyses and overview visualizations of the data set as presented by Klemm et al. \cite{Klemm2014VIS} are not tailored to a specific target variable and mostly highlight correlations between variables, which are known to the domain expert (e.g., correlation between body size and spine shape).
%%
We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis towards a specific disease or hypothesis.
%%
This is achieved by providing template regression formulas, which are applied to all potential variable combinations.
%%
Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
%%
Difference views between regression formulas allow to assess the influences of individual variables on the process.

Our contributions are:
\begin{itemize}
	\item An overview visualization technique representing feature interactions using target features.
	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
	\item The analysis of confounding variables by providing comparison views between different formula results.
%	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
	\item The expansion and adaption of our method to any data with our open and web-based system.
\end{itemize}
%%
The term Regression Cube has been used for the detailed analysis of one or a small number of regression models \cite{Chan, Ahmadi}.
%%
They extend 2D scatter plots to display detailed information about a regression model.
%%
In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as Regression Cube.
%%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features using statistical methods.
%%
As described by Thew et al. \cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{itemize}
	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of variables depicting the hypothesis and include confounding variables.
	\item Statisticians assess the association of the derived features towards the investigated disease.
\end{itemize}
%%
Relative risks can be determined if a statistical resilient association of features towards a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data using the classical sequential epidemiological workflow. Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

@@

The number of features calculated by the CFS algorithm is dependent on the information entropy in the data.
%%
In our epidemiological data, we usually observed a number of 10 to 20 features.
%%
% Was heißt der Termin in Klammern? 1. Weg; 2. "ca. 50% of the feature set 
For features, which affect the most other features (Recall "Confounder" in Section~\ref{sec:EpidemiologicalData}), e.g., age or gender,  the number is significantly larger (about half the features in the whole data set).
%%
The number of features extracted is not directly associated with the overall number, but rather on their information entropy towards each other.
%%
\com{Uli, is this correct this way?}
%%

@@

$Cancer \sim X + Y + Z + feature_1:feature_2$ encodes more assumptions.
%%
This formula models the hypothesis of an interaction between $feature_1$ and $feature_2$ (denoted with $:$) being relevant for the target feature, but it is not clear how other feature combination influence the result.
%%
Therefore, the cube incorporates this interaction for all $X$,$Y$ and $Z$ values as independent features.
%%

@@

\paragraph{3D-Prism as Data Mini-Map.}
\com{Add section about difference view for static target features.}
%%
The 3D-cube representation acts as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
%%
It serves an function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
%%
It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
%%
The displayed prism shows values above the matrix diagonal.
%%
%% TODO: Add reference for disadvantages of 3D information visualizations!
\paragraph{Tackling the disadvantages of 3D information visualization.}
3D-Information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
%%
We aim to minimize these problems as much as possible.
%%
Since the $R^2$ values are mapped on data point opacity, only large values are highlighted in the prism, guiding the focus to the respective slices.
%%
It also creates a very sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
%%
Also, the preceding correlation based feature selection reduces the information space significantly, leading to sparse cubes.
%%
Overlapping is still an issue, but this way greatly reduced in its affect to the readability of the visualization.
%%

%% INTERACTION
Rotating with the cube is restricted to the y-axis, preserving the mental map to position individual features.
%%
The cube is always oriented according to the 2D-representation, allowing for an easy mental combination of the two representations.
%%
Allowing more degrees of freedom was confusing to our users and also did not add value to the visualization.
%%
We provide also a zoom functionality using the mouse wheel input.
%%
%\\\\
\paragraph{Cube Slice Selection [3.ZF].}
%%
In order to \emph{Zoom, Filter and Analyze Further}, the user has to navigate towards different slices of interest.
%%
We propose two ways to achieve this.
\begin{itemize}
	\item \emph{Applying the slicing metaphor from 3D volume data.}
	%%
	In medical volume data renderings, slicing views very common to view details on a selected plane in the scene.
	%%
	We employ this technique for selecting cube slices (e.g. by moving a plane via vertical mouse input while pressing the right mouse button).
	%%
	We, however, still display the whole 3D-object instead of cutting away information towards the slice position.
	%%
	\item \emph{Selecting the slice using a dropdown menu} provides fast access to plane selections when the user already knows the slices of interest.
	%%
\end{itemize}
%%
The currently selected slice is denoted with a semi-transparent gray plane.
%%
The space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the current selected plane.
%%
the values are projected on this plane to give an overlapping-free view on the data points, which makes it easier to identify the current slice.
%%
%It also allows for a fast slicing through the cube and assessing the information.
%%
%\com{display of skewed cube, combination of 2D to 3D, navigating between the two views, projection of current slice to the 3D-Cube, details on demand}
%\\\\
\paragraph{2D Mosaic Plot Slice Visualization [4.DD].}
\com{Add Figure reference!}
%%
The 2D mosaic visualization shows all values below the matrix diagonal of the current slice, creating optical equivalence towards the 3D cube.
%%
To reduce visual clutter, the 2D view only shows dimensions, which are retrieved through the correlation based feature selection.
%%
The free space above the matrix diagonal is used to display the 3D cube.
%%

@@

%%
We conducted a study with two participants, who also co-authored this publication.
%%
Domain expert for the breast fat data set analysis is KH, a clinician (10 years of experience) with focus on epidemiological research.
%%
\emph{KH} is a radiologist, responsible for the SHIP-MRI acquisition and also responsible of the mammography analysis.

@@

\emph{TI}, a statistician with focus on epidemiology (\com{TODO} years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
%%
\subsection{The Breast Fat Data Set}
The breast fat data set was compiled to find associations between parenchyma tissue proportion in the female breast towards other features in the data.
%%
The ratio between parenchyma and cellular connective tissue (breast density) has been shown to be associated with breast cancer.
%%
Studies describe a four to five times increased risk of getting breast cancer for participants with a breast density above 50\% \cite{Mccormack2006}.

The data comprises 1.186 subjects (368 from \texttt{SHIP-2}, 818 from \texttt{SHIP-TREND-0} cohort).
%%
It contains 231 dimensions, holding information about
\begin{itemize}
	\item \emph{somatometric features}, e.g., body size, weight or BMI, 
	\item \emph{lifestyle features}, e.g., alcohol/tobacco consume, 
	\item \emph{personal history}, e.g., occupation, marital status,
	\item \emph{medical history}, e.g., current or prior diseases or laboratory values, such as cholesterol levels,
	\item \emph{women specific features}, e.g., number of born children, contraception type or hormone replacement therapy and
	\item \emph{mammography features} derived from the image data.
\end{itemize}
%%
The latter were derived using a level set based segmentation \cite{Ivanovska2014} of MRI image data for each subject.
%%
\com{Check this with Mrs. Hegenscheid, probably most of them were segmented by hand!}
%%
%\com{Describe Parenchym tissue and its relation to breast cancer.}
%\com{Describe how the data was acquired and preprocessed}

The data of each cohort was presented as individual \texttt{SPSS} files.
%%
All data related to the mammography was stored in a additional file.
%%
We converted the \texttt{SPSS} data sets to CSV and used \texttt{R} to merge the data sets together using their ID.
%%
All features were renamed to be expressive, e.g., \emph{chro\_09a} is now denoted as \emph{Disease\_Osteoporosis}.
%%
This avoids the need of defining a separate data dictionary file for the \emph{Regression Cube}, translating the feature names.
%%
All male subjects were removed as their data does not contribute to the analysis.
%%
\subsection{The Hepatic Steatosis Data Set}
We employ a subset of the data set which was used in \cite{Niemann2014} to identify predictive features with respect to the reversible disorder hepatic steatosis (fatty liver).
%%
The binary target feature is derived from the liver fat concentration which was measured with Magnetic Resonance Imaging (MRI).
%%
In particular, participants with a liver fat concentration of no more than 10\% are mapped to the 'negative' class; values greater than 10\% are mapped to the 'positive' class.
%%
The data set contains labels for 578 participants for which the MRI data was available to that point.
%%
Note that MRI was performed only in SHIP-2, so the target feature is present for just one of three study waves.

%%
Apart from the target feature, the data set contains 199 features extracted from participants' questionnaire answers and medical tests.
%%
They are features on socio-demographics (sex, age, etc.), features on consumption behavior (e.g. alcohol and cigarettes), SNPs, features extracted from laboratory data (e.g. sera concentrations), and two features on the results of the liver ultrasound.
%%
The names of 85 features have the suffix \emph{s0} meaning that their values have been recorded in SHIP-0 (first study moment), 50 features exhibit the suffix \emph{s1} and 55 features have the suffix \emph{s2}, alongside 10 time-independent SNPs.
%%

In \cite{Niemann2014}, the authors show that women and men exhibit differences in the class distribution of the liver fat concentration.
%%
Also, for women they identify that there is a association between age and liver fat concentration.
%%
They compute an appropriate cut-off value of 52 years for which the class distribution is most homogeneous within the resulting subsets.
%%
Based on this observations, we perform our analysis on three sub-populations, \emph{females}, \emph{males} and \emph{females} older than 52 years.
%%
\subsection{Case 1: Hypothesis-free Analysis of the Breast Cancer Data Set}
%%
%%
The analysis goal is to find relationships on the breast cancer data using features derived from the mammography analysis.
%%
Relationships between the share of parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
%%

\subsection{Case 2: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}
\com{Add R$^2$ values}
%%
We refer to each analysis step regarding to their belonging in the VA-Mantra (Sec.~\ref{sec:Workflow}).
%%
The analysis goal is reproducing results based on \cite{Niemann2014}.
%%
Therefore, \emph{UN} started the \textbf{AF} step using the dichotomized MRI fat liver concentration using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subjects.
%%
The \emph{SI} step using the 3D cube representations locates hotspots at the end of the cube (Fig.~\ref{fig:Application} a).
%%
\textbf{ZF} by slicing through the cube using the mouse input to inspect the hotspots using the mosaic plot (\textbf{DD}) revealed high correlations for somatometric features, hepatic steatosis indicator-features as well as laboratory values, such as creatinine (\com{important for kidney}) and uric acid (\com{gout, diabetes}) magnitudes.
%%
Similar results present for analyzing the \emph{female} groups.
%%
\emph{UN} could reproduce most results, some features features exhibit lower correlations though, e.g., creatinine magnitudes.
%%
Relationships not described in \cite{Niemann2014} were found, such as enzymes indicating liver dysfunctions, e.g. \emph{aspartate aminotransferase}.
%%
Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.

% %%
% The goal at the beginning of the analysis was to reproduce results based on the prior analysis.
% %%
% Subjects with a fatty liver could be identified using somatometric features, hepatic steatosis indicator-features as well as creatinine (\com{important for kidney}) and uric acid (\com{gout, diabetes}) magnitudes.
% %%
% The analysis was started with the target feature \emph{liver fat content} (\emph{mrt\_liverfat\_s2}), which was derived from the MRI data analysis.
% %%
% \emph{UN} therefore started the calculation using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subects.
% %%
% %%
% %%
% The data set containing all \emph{males} showed broad consensus with the expected results.
% %%
% The 3D-cube showed highlighted features at the end of the cube (Fig.~\ref{fig:application} a).
% %%
% Using the mouse-slicing allowed for fast navigation to the respective features.
% %%
% They include indicators hepatic steatosis ($R^2$:XX) as well as somatometric features, such as \emph{body mass index} (\emph{BMI}) and waist circumference.
% %%
% Feature correlations with \emph{gamma-glutamyl transpeptidase} (\emph{GGT}, indicating damaged livers if elevated) were found, but with low $R^2$ of XX.
%
% %%
% UN analyzed both \emph{female} groups in the same manner, concluding with the same findings regarding \emph{somatometric} features as well as \emph{hepatic steatosis} indicators.
% %%
% Using the mosaic plot, relationships incorporating enzyme used for the diagnosis of liver dysfunctions were found, e.g. aspartate aminotransferase.
% %%
%
% For most features, UN found at least tendencies using our approach.
% %%
% Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.
%
% %Regression cube calculations were carried out for

\textbf{Analysis of non-discretized target feature.}
%%
Since our method can analyze numerical target features, as opposed to decision trees used by \emph{UN} before, the analysis was conducted again for the not-dichotomized target using the same formula.
%%
The 3D-cube representation already showed generally lower $R^2$ values, but since the analysis is now based on linear regression, the $R^2$ values can not be directly compared.
%%
Correlation hot spots matched with the ones from the dichotomous target, but were generally lower ($R^2$ of XX for somatometric features as opposed to XX).
%%
We conclude that the bias introduced by dichotomizing the fat liver content enforces the findings towards liver diseases, while using the numerical features is less expressive.
%%

\textbf{Interleukin-6 correlation with liver fat.}
%%
During the analysis, one hotspot was always observed in the \emph{SI} and \emph{ZF} incorporating a high \emph{Interleukin-6} (\emph{IL-6}) correlation with liver fat values (Fig.~\ref{fig:Application} b).
%%
The correlation was high for both the dichotomized and continuous target feature.
%%
In the literature, relations between liver cancer \cite{He2013} and also chronic liver diseases \cite{Streetz2003}.
%%
Strong effects of \emph{IL-6} with hepatic steatosis were also prior described with mice \cite{Hong2004}.
%%
The finding is subject of further analysis.
%%
\subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
%%
Our method was well received among the domain experts.
%%
The novel way of modeling expert knowledge to derive hypothesis driven overview visualizations allow for detailed analysis of target features, either disease indicators or even image-derived features.
%%
%TI commented on the

>>>>

%\com{New Solution}.
%%
We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
%%
This approach combines all \emph{independent features} with respect to a \emph{target feature} and provides a visualization, which reveals insights into the data by highlighting relationships.
%%
A 3D-visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
%%
Slicing through the \emph{Regression Cube} allows for the detailed analysis of features towards the target disease.
%%
Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
%%
Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
%%
%\com{Validation \& Results}.
%%
We applied our \emph{Regression Cube} method to a hepatic steatosis data set to reproduce results from a data-mining driven analysis.
%%
A qualitative analysis was conducted on a breast fat data set.
%%
We were able to derive new hypotheses about relations between breast fat density and breast lesions towards breast cancer.
%%
%\com{Implications}.
%%
With the \emph{Regression Cube}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
} % end of abstract

@@

Testing features for association with diseases using regression analysis is one of the most important epidemiological tools.
%%
%Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the required subject count.
Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
%%
% missing overview techniques ist zu hart; es gibt ja techniken (eine scatterplot matrix zb.)
Due to the amount of data and only limited overview techniques, possible correlations may be missed.
%%
Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target variable and mostly highlight correlations between variables, which are known to the domain expert (e.g., correlation between body size and spine shape).
%%
We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis towards a specific disease or hypothesis.
%%
This is achieved by providing template regression formulas, which are applied to all potential variable combinations.
%%
Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
%%
Difference views between regression formulas allow to assess the influences of individual variables on the process.

Our contributions are:
\begin{itemize}
	\item An overview visualization technique representing feature interactions using target features.
	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
	\item The analysis of confounding variables by providing comparison views between different formula results.
%	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
	\item The expansion and adaption of our method to any data with our open web-based system.
\end{itemize}
%%
The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Chan, Ahmadi}.
%%
They extend 2D scatter plots to display detailed information about a regression model.
%%
In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as Regression Cube.
%%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features using statistical methods.
%%
As described by Thew et al. \cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{itemize}
	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of variables depicting the hypothesis and include confounding variables.
	\item Statisticians assess the association of the derived features towards the investigated disease.
\end{itemize}
%%
Relative risks can be determined if a statistical resilient association of features towards a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data using the classical sequential epidemiological workflow. Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

@@

The number of features calculated by the CFS algorithm is dependent on the information entropy in the data.
%%
In our epidemiological data, we usually observed a number of 10 to 30 features.
%For features, which affect the most other features (Recall "Confounder" in Section~\ref{sec:EpidemiologicalData}), e.g., age or gender,  the number is significantly larger (about half the features in the whole data set).
%%
The number of selected features using the CFS algorithm reflects their information entropy on the target.
%%
A large list of features is an expression of low correlation to the target feature.
%%
%The number of features extracted is not directly associated with the overall number, but rather on their information entropy towards each other.
%%

@@

$Cancer \sim X + Y + Z + feature_1:feature_2$ encodes more assumptions.
%%
This formula models the hypothesis of an interaction between $feature_1$ and $feature_2$ (denoted with `$:$') being relevant for the target feature, but it is not clear how other feature combination influence the result.
%%
Therefore, the cube incorporates this interaction for all $X$,$Y$ and $Z$ values as independent features.
%%

@@

\paragraph{3D-Prism as Data Mini-Map.}
%%
The 3D-cube representation acts as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
%%
It serves an function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
%%
%It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
%%
The displayed prism shows values above the matrix diagonal.
%%
For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes absolute $R^2$ values (Fig.~\ref{fig:System} b).
%%
Applying this strategy to a formula containing a static target (e.g. $Cancer \sim X + Y + Z$) yields many occlusions, since the CFS algorithm creates the same feature space for every slice.
%%
For such formulas, the 3D view encodes every data element as absolute difference between its $R^2$ features and the global mean along the z-axis.
%%
This highlights slices with unusually low or high results (Fig.~\ref{fig:Application}).
%%
%% TODO: Add reference for disadvantages of 3D information visualizations!
\paragraph{Tackling the disadvantages of 3D information visualization.}
3D-Information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
%%
We aim to minimize these problems as much as possible.
%%
Since the $R^2$ values are mapped on data point opacity, only large values are highlighted in the prism, guiding the focus to the respective slices.
%%
It also creates a very sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
%%
Also, the preceding correlation based feature selection reduces the information space significantly, leading to sparse cubes.
%%
Overlapping is still an issue, but this way greatly reduced in its affect to the readability of the visualization.
%%

%% INTERACTION
Rotating with the cube is restricted to the y-axis, preserving the mental map to position individual features.
%%
The cube is always oriented according to the 2D-representation, allowing for an easy mental combination of the two representations.
%%
Allowing more degrees of freedom was confusing to our users and also did not add value to the visualization.
%%
We provide also a zoom functionality using the mouse wheel input.
%%
%\\\\
\paragraph{Cube Slice Selection [3.ZF].}
%%
In order to \emph{Zoom, Filter and Analyze Further}, the user has to navigate towards different slices of interest.
%%
We propose two ways to achieve this.
\begin{itemize}
	\item \emph{Applying the slicing metaphor from 3D volume data.}
	%%
	In medical volume data renderings, slicing views very common to view details on a selected plane in the scene.
	%%
	We employ this technique for selecting cube slices (e.g. by moving a plane via vertical mouse input while pressing the right mouse button).
	%%
	We, however, still display the whole 3D-object instead of cutting away information towards the slice position.
	%%
	\item \emph{Selecting the slice using a dropdown menu} provides fast access to plane selections when the user already knows the slices of interest.
	%%
\end{itemize}
%%
The currently selected slice is denoted with a semi-transparent gray plane.
%%
The space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the current selected plane.
%%
the values are projected on this plane to give an overlapping-free view on the data points, which makes it easier to identify the current slice.
%%
%It also allows for a fast slicing through the cube and assessing the information.
%%
%\com{display of skewed cube, combination of 2D to 3D, navigating between the two views, projection of current slice to the 3D-Cube, details on demand}
%\\\\
\paragraph{2D Mosaic Plot Slice Visualization [4.DD].}
%%
The 2D mosaic visualization (Fig.~\ref{fig:System} c) shows all values below the matrix diagonal of the current slice, creating optical equivalence towards the 3D cube.
%%
To reduce visual clutter, the 2D view only shows dimensions, which are retrieved through the correlation based feature selection.
%%
The free space above the matrix diagonal is used to display the 3D cube.
%%

@@

%%
We conducted a study with two participants, who also co-authored this publication.
%%
Domain expert for the breast fat data set analysis is \emph{KH}, a clinician (10 years of experience) with focus on epidemiological research.
%%
\emph{KH} is a radiologist, responsible for the SHIP-MRI acquisition and also responsible of the mammography analysis.

@@

\emph{TI}, a statistician with focus on epidemiology (8 years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
%%
\subsection{The Hepatic Steatosis Data Set}
We employ the data set used in \cite{Niemann2014} to identify predictive features w.r.t. the reversible hepatic steatosis disorder.
%%
The dichotomous target feature is derived from the liver fat concentration measured using MRI scans.
%%
Liver fat concentrations of no more than 10\% are mapped to the 'negative' class; values greater than 10\% are mapped to the 'positive' class.
%%
The data set contains labels for 578 participants for which the MRI data was available to that point.
%%
The MRI scans for each subject were introduced in SHIP-2, so the target feature is present for just one of three study waves.

%%
Apart from the target feature, the data set contains 199 features, comprising socio-demographic features (gender, age), consumption behavior (e.g. alcohol or tobacco), laboratory data (e.g. sera concentrations), and two features depicting the liver ultrasound.
%%
%The acquisition wave is denoted using the appendix; 85 for S0 (SHIP-0), 55 for S1, 55 for S2 and 10 time-independent SNPs (DNA base pairs).
%%
The acquisition wave is denoted using the appendix; \emph{s0} meaning that their values have been recorded in SHIP-0 (first study moment), 50 features for \emph{s1} and 55 for \emph{s2}, alongside with 10 time-independent SNPs (DNA base pairs).
%%

In \cite{Niemann2014}, the authors show different class distribution of liver fat concentrations of women and men.
%%
For women a association between age and liver fat was identified.
%Also, for women they identify that there is a association between age and liver fat concentration.
%%
An appropriate cut-off value of 52 years was set, yielding the most homogeneous class distribution within the resulting subsets.
%%
Based on these observations, we perform our analysis on three sub-populations: \emph{females}, \emph{males} and \emph{females} older than 52 years.
%%
\subsection{The Breast Fat Data Set}
The breast fat data set was compiled to find associations between parenchyma tissue proportion in the female breast towards other features in the data.
%%
The ratio between parenchyma and cellular connective tissue (breast density) has been shown to be associated with breast cancer.
%%
Studies describe a four to five times increased risk of getting breast cancer for participants with a breast density above 50\% \cite{Mccormack2006}.

The data comprises 1.186 female subjects (368 from \texttt{SHIP-2}, 818 from \texttt{SHIP-TREND-0} cohort).
%%
It contains 231 dimensions, holding information about somatometric features (e.g. body size or weight) consumption behavior (e.g. alcohol or tobacco), personal and medical history (e.g., occupation or prior diseases), women specific features (e.g., number of born children or contraception type) as well as mammography features.
% It contains 231 dimensions, holding information about
% \begin{itemize}
% 	\item \emph{somatometric features}, e.g., body size, weight or BMI,
% 	\item \emph{lifestyle features}, e.g., alcohol/tobacco consume,
% 	\item \emph{personal history}, e.g., occupation, marital status,
% 	\item \emph{medical history}, e.g., current or prior diseases or laboratory values, such as cholesterol levels,
% 	\item \emph{women specific features}, e.g., number of born children, contraception type or hormone replacement therapy and
% 	\item \emph{mammography features} derived from the image data.
% \end{itemize}
%%
The latter were derived using manual segmentation of MRI image data for each subject carried out by radiologists.
%%
%%
%\com{Describe Parenchym tissue and its relation to breast cancer.}
%\com{Describe how the data was acquired and preprocessed}

The data of each cohort was presented as individual \texttt{SPSS} files.
All data related to the mammography was stored in a additional file.
We converted the \texttt{SPSS} data sets to CSV and used \texttt{R} to merge the data sets together using their ID.
All features were renamed to be expressive, e.g., \emph{chro\_09a} is now denoted as \emph{Disease\_Osteoporosis}.
This avoids the need of defining a separate data dictionary file for translating the feature names.
%All male subjects were removed as their data does not contribute to the analysis.
\subsection{Case 1: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}
We refer to each analysis step regarding to their belonging in the VA-Mantra (Sec.~\ref{sec:Workflow}).
The analysis goal is reproducing results based on \cite{Niemann2014}.
Therefore, \emph{UN} started the \textbf{1.AF} step using the dichotomized MRI fat liver concentration using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subjects.
The \textbf{2.SI} step using the 3D cube representations locates hotspots at the end of the cube (Fig.~\ref{fig:Application} a).
\textbf{3.ZF} by slicing through the cube using the mouse input to inspect the hotspots using the mosaic plot (\textbf{4.DD}) revealed high correlations for somatometric features, hepatic steatosis indicator-features as well as laboratory values, such as creatinine (used renal retention parameter) and uric acid (among others used as gout and diabetes risk factors) magnitudes.
Similar results present for analyzing the \emph{female} groups.
\emph{UN} could reproduce most results, some features features exhibit lower correlations though, e.g., creatinine magnitudes.
A slight influence of age on the target feature could be observed for women ($R^2$ of $0.09$ for females compared to $0.02$ for males).
Relationships not described in \cite{Niemann2014} were found, such as enzymes indicating liver dysfunctions, e.g. \emph{aspartate aminotransferase}.
Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.
% %%
% The goal at the beginning of the analysis was to reproduce results based on the prior analysis.
% %%
% Subjects with a fatty liver could be identified using somatometric features, hepatic steatosis indicator-features as well as creatinine (\com{important for kidney}) and uric acid (\com{gout, diabetes}) magnitudes.
% %%
% The analysis was started with the target feature \emph{liver fat content} (\emph{mrt\_liverfat\_s2}), which was derived from the MRI data analysis.
% %%
% \emph{UN} therefore started the calculation using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subects.
% %%
% %%
% %%
% The data set containing all \emph{males} showed broad consensus with the expected results.
% %%
% The 3D-cube showed highlighted features at the end of the cube (Fig.~\ref{fig:application} a).
% %%
% Using the mouse-slicing allowed for fast navigation to the respective features.
% %%
% They include indicators hepatic steatosis ($R^2$:XX) as well as somatometric features, such as \emph{body mass index} (\emph{BMI}) and waist circumference.
% %%
% Feature correlations with \emph{gamma-glutamyl transpeptidase} (\emph{GGT}, indicating damaged livers if elevated) were found, but with low $R^2$ of XX.
%
% %%
% UN analyzed both \emph{female} groups in the same manner, concluding with the same findings regarding \emph{somatometric} features as well as \emph{hepatic steatosis} indicators.
% %%
% Using the mosaic plot, relationships incorporating enzyme used for the diagnosis of liver dysfunctions were found, e.g. aspartate aminotransferase.
% %%
%
% For most features, UN found at least tendencies using our approach.
% %%
% Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.
%
% %Regression cube calculations were carried out for

\textbf{Analysis of non-discretized target feature.}
%%
Since our method can analyze numerical target features, as opposed to decision trees used by \emph{UN} before, the analysis was conducted again for the not-dichotomized target using the same formula.
%%
The 3D-cube representation already showed generally lower $R^2$ values, but since the analysis is now based on linear regression, the $R^2$ values can not be directly compared.
%%
Correlation hot spots matched with the ones from the dichotomous target, but were generally lower ($R^2$ of $0.37$ for somatometric features as opposed to $0.58$).
%%
We conclude that the bias introduced by dichotomizing the fat liver content enforces the findings towards liver diseases, while using the numerical features is less expressive.
%%

\textbf{Interleukin-6 correlation with liver fat.}
%%
During the analysis, one hotspot was always observed in the \emph{SI} and \emph{ZF} incorporating a high \emph{Interleukin-6} (\emph{IL-6}) correlation with liver fat values (Fig.~\ref{fig:Application} b).
%%
The correlation was high for both the dichotomized and continuous target feature.
%%
In the literature, relations between liver cancer \cite{He2013} and also chronic liver diseases \cite{Streetz2003}.
%%
Strong effects of \emph{IL-6} with hepatic steatosis were also prior described with mice \cite{Hong2004}.
%%
The finding is subject of further analysis.
%%
\subsection{Case 2: Hypothesis-free Analysis of the Breast Cancer Data Set}
%%
The analysis goal is to find relationships on the breast cancer data using features derived from the mammography analysis.
%%
Relationships between the share of parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
%%
\subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
%%
Our method was well received among the domain experts.
%%
For the first time, they were able to derive a overview visualization custom tailored to underlying assumptions.
%%
In this context, \emph{TI} highlighted the ability to simultaneously analyze thousands of regression models while maintaining little time expenses for rating them.
%The novel way of modeling knowledge to derive hypothesis driven overview visualizations allow for detailed analysis of target features, either disease indicators or even image-derived features.
%%
%TI commented on the
